---
title: "Standalone Queries using rOpenSci ‘traits' package"
author: "David LeBauer"
date: "12/10/2018"
output:
  pdf_document: default
  html_document: default
---

The objective of this document is to demonstrate how to access the yield data and agronomic metadata used to produce figure 4 of the manuscript. This repeats the queries done in the gcbb_manuscript.Rmd file, except that it uses the Ropensci traits package to query the betydb.org API. Therefore this example does _not_ require direct access to the database. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup


```{r packages}
if(!require(traits)) install.packages('traits')
library(traits)
library(dplyr)
library(ggplot2)

options(betydb_url = 'http://betydb.org',
        betydb_api_version = 'v1',
        betydb_key = 'Kq2xEWiRCyj90Sq7EFVn2A0LxJb58UTbscJl4fJP'
        # key is for guest user; 
        # provides access to all public data 
        # (this includes all data used in the ma)
        ) 
```



### Query Yield data for switchgrass and miscanthus


```{r yield-query, cache = TRUE}
yields <- betydb_search(result_type = 'yields', limit = 'none')

grass_yields <- yields %>% 
    dplyr::filter(genus %in% c('Miscanthus', 'Panicum')) 

```




## Query and agronomic metadata 

treatments are categorical, each study has >=1 treatment; managements describe the actual activities (planting, fertilization, irrigation, etc) and sometimes the level (planting density, fertilization rate, etc). 

There is a many-to-many relationship between treatments and managements. One treatment can have many managements (e.g. control treatment had a planting date, a level of fertilization, etc). And each management can be associated with one or more treatments - e.g. the same planting for both a control and fertilized treatment.

So first we query the tables, then join them, then create new columns for the date and level of specific managements.


```{r managements, cache=TRUE}
treatments <- betydb_query(table = 'treatments', limit = 'none') %>% 
  dplyr::mutate(treatment_id = id) %>% 
  dplyr::select(treatment_id, name, definition, control)

managements <- betydb_query(table = 'managements', limit = 'none') %>%
  dplyr::filter(mgmttype %in% c('fertilizer_N', 'fertilizer_N_rate', 'planting', 'irrigation')) %>%
  dplyr::mutate(management_id = id) %>%
  dplyr::select(management_id, date, mgmttype, level, units) 

# now link managements to treatments
m <- betydb_query(table = 'managements', associations_mode = 'ids', limit = 'none') 

managements_treatments <- m %>%
 select(treatment_id = `associated treatment ids`, management_id = id) %>% 
  tidyr::unnest()

managements <- managements %>%
  left_join(managements_treatments, by = 'management_id') %>%
  left_join(treatments, by = 'treatment_id') 
```

## Now compute specific managements of interest

```{r management_events}
nitrogen <- managements %>% 
  dplyr::filter(mgmttype == "fertilizer_N_rate") %>%
  dplyr::select(treatment_id, nrate = level)
planting <- managements %>% 
  dplyr::filter(mgmttype == "planting") %>%
  dplyr::select(treatment_id, planting_date = date)
planting_rate <- managements %>% 
  dplyr::filter(mgmttype == "planting") %>%
  dplyr::select(treatment_id, planting_date = date, planting_density = level) %>% 
  dplyr::filter(!is.na(planting_density))
irrigation <- managements %>% 
  dplyr::filter(mgmttype == 'irrigation') 
irrigation_rate <- irrigation %>% 
  dplyr::filter(units == 'mm', !is.na(treatment_id)) %>% 
  group_by(treatment_id, year = sql("extract(year from date)"), units) %>% 
  summarise(irrig.mm = sum(level)) %>% 
  group_by(treatment_id) %>% 
  summarise(irrig.mm.y = mean(irrig.mm))
irrigation_boolean <- irrigation %>%
  group_by(treatment_id) %>% 
  dplyr::mutate(irrig = as.logical(mean(level))) %>% 
  dplyr::select(treatment_id, irrig = irrig)
irrigation_all <- irrigation_boolean %>%
  full_join(irrigation_rate, copy = TRUE, by = 'treatment_id')
```

### Subset grasses; combine with agronomic metadata

Now lets join the agronomic metadata to the yield data:

```{r}
grass_yields <- grass_yields %>% 
  dplyr::filter(genus %in% c('Miscanthus', 'Panicum')) %>%
  left_join(nitrogen, by = 'treatment_id') %>% 
  #left_join(planting, by = 'treatment_id') %>% 
  left_join(planting_rate, by = 'treatment_id') %>% 
  left_join(irrigation_all, by = 'treatment_id', copy = TRUE) %>% 
  dplyr::mutate(age = lubridate::year(raw_date)- lubridate::year(planting_date),
         nrate = ifelse(is.na(nrate), 0, nrate),
         SE = ifelse(statname == "SE", stat, ifelse(statname == 'SD', stat / sqrt(n), NA)),
         continent = ifelse(lon < -30, 'united_states', ifelse(lon < 75, 'europe', 'asia'))) %>%
  dplyr::select(date, lat, lon, nrate, planting_date, planting_density, irrig, 
         irrig.mm.y, age, mean, n, SE, scientificname, genus, continent, 
         sitename, author, year) %>% 
  dplyr::filter(!duplicated(.))
save(grass_yields, file = "grass_yields.RData")

```

### Reproduce figure 4a (but without regression fits)

We are just focusing on fig 4a because the scope of this is to demonstrate access to the yield data in BETYdb. The code required to fit regressions and download / prepare met data in the original analysis (gcbb_manuscript.Rmd) should still work.

```{r}
ggplot(data = grass_yields, aes(x = nrate, color = genus)) +
  geom_point(aes(x = jitter(nrate, 20), y = mean), alpha = 0.25, size = 0.25) +
  ylab(expression(Yield~~"(Mg "*ha^"-1"*yr^"-1"*")")) +
  xlab(expression("Nitrogen Fertilization Rate"~~"(kg "*ha^"-1"*yr^"-1"*")")) + 
  xlim(0,250) +
  scale_colour_brewer(palette = "Set1", labels = c('Miscanthus', 'Panicum (Switchgrass)'))
```
